Accelerating eigenvector and pseudospectra computation using blocked multi-shift triangular solves

### 2016-07-21 mstrsmPaper.tex
Adding appendix on generalized multi-shift triangular solve

  \end{algorithmic}
\end{algorithm}
\noindent
Applying this algorithm to one right-hand side takes \(m\) divisions
in the diagonal step and \(\sim m^2 \) flops in the substitution step,
so the computational work is dominated by the substitution step if the
matrix dimension is sufficiently large. This routine is backward
stable and implemented in the Level 2 BLAS routine
\texttt{xTRSV}. However, given a block size \(n_b\), we can apply a
blocked algorithm \cite{golub2012matrix}:
\begin{algorithm}[H]
  \label{algorithm:trsm}
  \caption{Triangular solve with blocked back substitution}
  \begin{algorithmic}
    \Procedure{Trsm}{$U,B$}
    \Comment \(B\) is overwritten with \(X\)

@@

\paragraph{}
The multi-shift triangular solve problem is a variant of the standard
triangular solve problem. Given scalar shifts
\(\lambda_1,\cdots,\lambda_n\),
we seek to solve \(n\) triangular systems of the form
\begin{align}
  \left(U-\lambda_j I\right) x_j &= b_j.
\end{align}
Each triangular system has a different matrix, so a naive approach is
to apply unblocked back substitution on each system. However, the
matrices only differ in the diagonal entries. This implies we can use
blocked back substitution with a modification to the diagonal block
step:
\begin{algorithm}[H]
  \label{algorithm:multi-shift trsm}
  \caption{Multi-shift triangular solve with blocked back substitution}
  \begin{algorithmic}
    \Procedure{MultiShiftTrsm}{$U, \lambda, B $}
    \Comment \(B\) is overwritten with \(X\)

@@

    \State \(b := t * b\)

@@

    \State \(b := t *b\)

@@

\paragraph{} Safeguarded back substitution is easily generalized to
the safe multi-shift triangular solve problem. Given nonzero
right-hand side vectors \(b_1,\cdots,b_n\) and shifts
\(\lambda_1,\cdots,\lambda_n\), we seek nonzero solution vectors
\(x_1,\cdots,x_n\) and scaling factors \(s_1,\cdots,s_n\) in
\(\left[0,1\right]\) such that
\begin{align}
  \left( U - \lambda_j I\right) x_j &= s_j b_j.
\end{align}
For safeguarded, blocked back substitution, we just need to ensure
that the matrix multiplications in the substitution step do not cause
numerical issues:
\begin{algorithm}[H]
  \label{algorithm:safe multi-shift trsm}
  \caption{Safe multi-shift triangular solve with safeguarded, blocked back substitution }
  \begin{algorithmic}
    \Procedure{SafeMultiShiftTrsm}{$U, \lambda, B, s $}
    \Comment \(B\) is overwritten with \(X\)

@@

    \State \(B(\mathcal{I}_0,j) := t*B(\mathcal{I}_0,j)\)

    \State \(B(\mathcal{I}_2,j) := t*B(\mathcal{I}_2,j)\)

@@

    \State \(B(:,j) := t * B(:,j)\)

@@

\section{Generalized Multi-Shift Triangular Solve}
We introduce a final, generalized variant of the multi-shift
triangular solve problem. Given \(m\times m\) upper triangular
matrices \(U\) and \(V\), right-hand side vectors \(b_1,\cdots,b_n\),
and shifts \(\lambda_1,\cdots,\lambda_n\), we seek nonzero solution
vectors \(x_1,\cdots,x_n\) such that
\begin{align}
  \left( U - \lambda_j V \right) x_j &= b_j .
\end{align}
This problem can be solved with a slight modification of Algorithm
\ref{algorithm:multi-shift trsm}:
\begin{algorithm}[H]
  \label{algorithm:generalized multi-shift trsm}
  \caption{Generalized multi-shift triangular solve with blocked back substitution}
  \begin{algorithmic}
    \Procedure{GeneralizedMultiShiftTrsm}{$U, V, \lambda, B $}
    \Comment \(B\) is overwritten with \(X\)

    \State \Call{Trsv}{$U', B(\mathcal{I}_1,j)$} 
    \Comment Diagonal block step (\texttt{xTRSV})

    \EndProcedure
  \end{algorithmic}
\end{algorithm}
\noindent

\section{Eigenvector Computation} \label{section:eigenvector}
We shall now apply a safe multi-shift triangular solve to compute the
eigenvectors of a general \(n\times n\) matrix \(A\).  Assuming that
\(A\) is nondefective, we seek an eigenvalue decomposition
\(A=X\Lambda X^{-1}\) where \(\Lambda\) is a diagonal eigenvalue
matrix and \(X\) an eigenvector matrix.\footnote{ If \(A\) is
  defective, a nondefective matrix can be obtained with a small
  perturbation of \(A\).}  We begin by computing the Schur
decomposition \(A=Q TQ^H\) where \(T\) is upper triangular and \(Q\)
unitary. In LAPACK's general eigensolver routine \texttt{xGEEV}, this
is performed efficiently with Level 3 BLAS by converting \(A\) to
upper Hessenberg form (\texttt{xGEHRD}) \cite{quintana2006improving},
converting Householder transforms to a unitary matrix
(\texttt{xUNGHR}), and applying the QR algorithm (\texttt{xHSEQR})
\cite{braman2002multishift1, braman2002multishift2}. Since similar
matrices have identical eigenvalues, \(\Lambda\) can be obtained by
simply reading off the diagonal of \(T\).  Now, all that remains is to
find a triangular eigenvector matrix \(Z\) such that \(T=Z \Lambda
Z^{-1}\) and to compute the back substitution \(X=QZ\).  Assuming that
\(Z\) is upper triangular and letting \(\lambda_k\) and \(z_k\)
respectively denote the \(k\)th eigenvalue and triangular eigenvector,
we require
\begin{align}
  \left[
  \begin{matrix}
    T_{11} & u & T_{13} \\
    0 & \lambda_k & v^T \\
    0 & 0 & T_{33}
  \end{matrix}
  \right]
  \left[
  \begin{matrix}
    \hat{z} \\ s \\ 0
  \end{matrix}
  \right]
  &=
  \lambda_k
  \left[
  \begin{matrix}
    \hat{z} \\ s \\ 0
  \end{matrix}
  \right].
\end{align}
This system is satisfied if and only if \(\hat{z}\) is a solution to a
\(k-1\times k-1\) shifted triangular system,
\begin{align}
  \left(T_{11} - \lambda_k I\right) \hat{z} &= -s u.
  \label{eq:triangeig-2}
\end{align}
This is similar to the form of the safe multi-shift triangular solve
problem, but each triangular system has a different size. LAPACK's
triangular eigenvector routine \texttt{xTREVC} approaches this problem
by calling \texttt{xLATRS} for each triangular eigenvector and back
transforming with calls to \texttt{xGEMV}. These routines are limited
to Level 2 BLAS and hence achieve poor performance. On multicore
architectures, this procedure can be accelerated by performing
\texttt{xLATRS } in parallel and by blocking the back substitution
into Level 3 BLAS \texttt{xGEMM} calls
\cite{gates2014accelerating}. However, a hardware-independent solution
is preferable for the sake of portability.  Observe that
(\ref{eq:triangeig-2}) holds if and only if
\begin{align}
  \left(\left[
    \begin{matrix}
      T_{11} & u & T_{13} \\
      0 & \lambda_k & v^T \\
      0 & 0 & T_{33}
    \end{matrix}
  \right] - \lambda_k I\right)
  \left[
    \begin{matrix}
      \hat{z} \\ 0 \\ 0
    \end{matrix}
  \right]
  &=
  s \left[
    \begin{matrix}
      -u \\ 0 \\ 0
    \end{matrix}
  \right].
\end{align}
This is the form for a safe multi-shift triangular solve.\footnote{If
  we apply back substitution to a right-hand vector where the last
  \(n-k+1\) entries are zero, the corresponding entries in the
  solution will also be zero.} Furthermore, the right-hand side matrix
will be strictly upper triangular, so we can shortcut the back
substitution to avoid unnecessary flops involving zeros. After
computing a safe multi-shift triangular solve, the triangular
eigenvectors are obtained by putting the scaling factors on the
diagonal of the solution matrix. The algorithm is outlined below:
\begin{algorithm}[H]
  \caption{Triangular and general eigensolver}
  \label{algorithm:eig}
  \begin{algorithmic}

@@

\paragraph{} Pseudospectra are typically computed by establishing a
grid with \(N\) points on a region of the complex plane, computing the
resolvent norm \(\lVert (zI - A)^{-1} \rVert_2\) at each grid point
\(z\), and visualizing with a contour plotter. Letting
\(\sigma_\text{max}(\cdot)\) and \(\sigma_\text{min}(\cdot)\) denote
the largest and smallest singular values of an input matrix,
respectively, we remark that the resolvent norm satisfies
\begin{align}
  \left\lVert \left(zI - A\right)^{-1} \right\rVert_2
  &= \sigma_\text{max}\left( \left(zI - A\right)^{-1} \right) \nonumber \\
  &= \frac{1}{\sigma_\text{min}\left( zI - A \right)}.
\end{align}
Thus, one could naively compute pseudospectra by computing the SVD of
\(zI-A\)
for each grid point \(z\)
and reporting the reciprocal of the smallest singular value. However,
this involves a total of \(O(Nn^3)\)
flops, which is prohibitively expensive for large matrices unless the
grid is very coarse. The Van Loan/Lui algorithm improves on the
computational cost by proceeding in two stages \cite{van1984near,
  lui1997computation}.\footnote{Most authors attribute this algorithm
  to Lui, but the algorithm was presented (in a different context) by
  Van Loan more than a decade earlier.} It begins by computing a Schur
decomposition \(A=QTQ^H\)
where \(T\)
is upper triangular and \(Q\)
unitary (see discussion in Section \ref{section:eigenvector}). Since
matrix norms are invariant under unitary transformations,
\begin{align*}
  \left\lVert \left( zI-A \right)^{-1} \right\rVert_2
  &= \left\lVert \left( zQQ^H - QTQ^H \right)^{-1} \right\rVert_2 \\
  &= \left\lVert -Q\left( T -zI\right)^{-1} Q^H \right\rVert_2 \\
  &= \left\lVert \left( T-zI \right)^{-1}\right\rVert_2 \\
  &= \sigma_\text{max}\left( \left( T-zI \right)^{-1} \right).
\end{align*}
Letting \(\lambda_\text{max}(\cdot)\) denote the largest eigenvalue of
a Hermitian matrix,
\begin{align}
  \left\lVert \left( zI-A \right)^{-1} \right\rVert_2
  &= \sqrt{\lambda_\text{max}\left( \left( T-zI \right)^{-H} \left( T-zI \right)^{-1} \right)}.
\end{align}
A Krylov eigensolver like the Lanczos algorithm can estimate the
largest eigenvalue of \(\left( T-zI \right)^{-H} \left( T-zI
\right)^{-1}\) by repeatedly applying it to a vector. Each matrix
product can be computed with two triangular solves, for a total cost
of \(O(n^2)\) flops. The algorithm is outlined below:
\begin{algorithm}[H]
  \label{algorithm:van loan/lui}
  \caption{Van Loan/Lui algorithm}
  \begin{algorithmic}
    \Function{VanLoanLui}{$A, z$}

@@

>>>>

  \end{algorithmic}
\end{algorithm}
\noindent
Applying this algorithm to one right-hand side takes \(m\) divisions
in the diagonal step and \(\sim m^2 \) flops in the substitution step,
so the computational work is dominated by the substitution step if the
matrix dimension is sufficiently large. This routine is backward
stable and implemented in the Level 2 BLAS routine
\texttt{xTRSV}. However, given a block size \(n_b\), we can apply a
blocked algorithm \cite{golub2012matrix}:
\begin{algorithm}[H]
  \caption{Triangular solve with blocked back substitution}
  \label{algorithm:trsm}
  \begin{algorithmic}
    \Procedure{Trsm}{$U,B$}
    \Comment \(B\) is overwritten with \(X\)

@@

\paragraph{}
The multi-shift triangular solve problem is a variant of the standard
triangular solve problem. Given scalar shifts
\(\lambda_1,\cdots,\lambda_n\),
we seek to solve \(n\) triangular systems of the form
\begin{align}
  \left(U-\lambda_j I\right) x_j &= b_j.
\end{align}
Each triangular system has a different matrix, so a naive approach is
to apply unblocked back substitution on each system. However, the
matrices only differ in the diagonal entries. This implies we can use
blocked back substitution with a modification to the diagonal block
step:
\begin{algorithm}[H]
  \caption{Multi-shift triangular solve with blocked back substitution}
  \label{algorithm:multi-shift trsm}
  \begin{algorithmic}
    \Procedure{MultiShiftTrsm}{$U, \lambda, B $}
    \Comment \(B\) is overwritten with \(X\)

@@

    \State \(b := t * b\) \Comment \texttt{xSCAL}

@@

    \State \(b := t *b\) \Comment \texttt{xSCAL}

@@

\paragraph{} Safeguarded back substitution is easily generalized to
the safe multi-shift triangular solve problem. Given nonzero
right-hand side vectors \(b_1,\cdots,b_n\) and shifts
\(\lambda_1,\cdots,\lambda_n\), we seek nonzero solution vectors
\(x_1,\cdots,x_n\) and scaling factors \(s_1,\cdots,s_n\) in
\(\left[0,1\right]\) such that
\begin{align}
  \left( U - \lambda_j I\right) x_j &= s_j b_j.
\end{align}
For safeguarded, blocked back substitution, we just need to ensure
that the matrix multiplications in the substitution step do not cause
numerical issues:
\begin{algorithm}[H]
  \caption{Safe multi-shift triangular solve with safeguarded, blocked back substitution }
  \label{algorithm:safe multi-shift trsm}
  \begin{algorithmic}
    \Procedure{SafeMultiShiftTrsm}{$U, \lambda, B, s $}
    \Comment \(B\) is overwritten with \(X\)

@@

    \State \(B(\mathcal{I}_0,j) := t*B(\mathcal{I}_0,j)\) \Comment \texttt{xSCAL}

    \State \(B(\mathcal{I}_2,j) := t*B(\mathcal{I}_2,j)\) \Comment \texttt{xSCAL}

@@

    \State \(B(:,j) := t * B(:,j)\) \Comment \texttt{xSCAL}

@@

\section{Eigenvector Computation} \label{section:eigenvector}
We shall now apply a safe multi-shift triangular solve to compute the
eigenvectors of a general \(n\times n\) matrix \(A\).  Assuming that
\(A\) is nondefective, we seek an eigenvalue decomposition
\(A=X\Lambda X^{-1}\) where \(\Lambda\) is a diagonal eigenvalue
matrix and \(X\) an eigenvector matrix.\footnote{ If \(A\) is
  defective, a nondefective matrix can be obtained with a small
  perturbation of \(A\).}  We begin by computing the Schur
decomposition \(A=Q TQ^H\) where \(T\) is upper triangular and \(Q\)
unitary. In LAPACK's general eigensolver routine \texttt{xGEEV}, this
is performed efficiently with Level 3 BLAS by converting \(A\) to
upper Hessenberg form (\texttt{xGEHRD}) \cite{quintana2006improving},
converting Householder transforms to a unitary matrix
(\texttt{xUNGHR}), and applying the QR algorithm (\texttt{xHSEQR})
\cite{braman2002multishift1, braman2002multishift2}. Since similar
matrices have identical eigenvalues, \(\Lambda\) can be obtained by
simply reading off the diagonal of \(T\).  Now, all that remains is to
find a triangular eigenvector matrix \(Z\) such that \(T=Z \Lambda
Z^{-1}\) and to compute the back substitution \(X=QZ\).  Assuming that
\(Z\) is upper triangular and letting \(\lambda_k\) and \(z_k\)
respectively denote the \(k\)th eigenvalue and triangular eigenvector,
we require
\begin{align}
  \left[
  \begin{matrix}
    T_{11} & u & T_{13} \\
    0 & \lambda_k & v^T \\
    0 & 0 & T_{33}
  \end{matrix}
  \right]
  \left[
  \begin{matrix}
    \hat{z} \\ s \\ 0
  \end{matrix}
  \right]
  &=
  \lambda_k
  \left[
  \begin{matrix}
    \hat{z} \\ s \\ 0
  \end{matrix}
  \right].
\end{align}
This system is satisfied if and only if \(\hat{z}\) is a solution to a
\(k-1\times k-1\) shifted triangular system,
\begin{align}
  \left(T_{11} - \lambda_k I\right) \hat{z} &= -s u.
  \label{eq:triangeig-2}
\end{align}
This is similar to the form of the safe multi-shift triangular solve
problem, but each triangular system has a different size. LAPACK's
triangular eigenvector routine \texttt{xTREVC} approaches this problem
by calling \texttt{xLATRS} for each triangular eigenvector and back
transforming with calls to \texttt{xGEMV}. These routines are limited
to Level 2 BLAS and hence achieve poor performance. On multicore
architectures, this procedure can be accelerated by performing
\texttt{xLATRS } in parallel and by blocking the back substitution
into Level 3 BLAS \texttt{xGEMM} calls
\cite{gates2014accelerating}. However, a hardware-independent solution
is preferable for the sake of portability.  Observe that
(\ref{eq:triangeig-2}) holds if and only if
\begin{align}
  \left(\left[
    \begin{matrix}
      T_{11} & u & T_{13} \\
      0 & \lambda_k & v^T \\
      0 & 0 & T_{33}
    \end{matrix}
  \right] - \lambda_k I\right)
  \left[
    \begin{matrix}
      \hat{z} \\ 0 \\ 0
    \end{matrix}
  \right]
  &=
  s \left[
    \begin{matrix}
      -u \\ 0 \\ 0
    \end{matrix}
  \right].
\end{align}
This is the form for a safe multi-shift triangular solve.\footnote{If
  we apply back substitution to a right-hand vector where the last
  \(n-k+1\) entries are zero, the corresponding entries in the
  solution will also be zero.} Furthermore, the right-hand side matrix
will be strictly upper triangular, so we can shortcut the back
substitution to avoid unnecessary flops involving zeros. After
computing a safe multi-shift triangular solve, the triangular
eigenvectors are obtained by putting the scaling factors on the
diagonal of the solution matrix. The algorithm is outlined below:
\begin{algorithm}[H]
  \caption{Triangular and general eigensolvers}
  \label{algorithm:eig}
  \begin{algorithmic}

@@

\paragraph{} Pseudospectra are typically computed by establishing a
grid with \(N\) points on a region of the complex plane, computing the
resolvent norm \(\lVert (zI - A)^{-1} \rVert_2\) at each grid point
\(z\), and visualizing with a contour plotter. Letting
\(\sigma_\text{max}(\cdot)\) and \(\sigma_\text{min}(\cdot)\) denote
the largest and smallest singular values of an input matrix,
respectively, we remark that the resolvent norm satisfies
\begin{align}
  \left\lVert \left(zI - A\right)^{-1} \right\rVert_2
  &= \sigma_\text{max}\left( \left(zI - A\right)^{-1} \right) \nonumber \\
  &= \frac{1}{\sigma_\text{min}\left( zI - A \right)}.
\end{align}
Thus, one could naively compute pseudospectra by computing the SVD of
\(zI-A\)
for each grid point \(z\)
and reporting the reciprocal of the smallest singular value. However,
this involves a total of \(O(Nn^3)\)
flops, which is prohibitively expensive for large matrices unless the
grid is very coarse. The Van Loan/Lui algorithm improves on the
computational cost by proceeding in two stages \cite{van1984near,
  lui1997computation}.\footnote{Most authors attribute this algorithm
  to Lui, but the algorithm was presented (in a different context) by
  Van Loan more than a decade earlier.} It begins by computing a Schur
decomposition \(A=QTQ^H\)
where \(T\)
is upper triangular and \(Q\)
unitary (see discussion in Section \ref{section:eigenvector}). Since
matrix norms are invariant under unitary transformations,
\begin{align*}
  \left\lVert \left( zI-A \right)^{-1} \right\rVert_2
  &= \left\lVert \left( zQQ^H - QTQ^H \right)^{-1} \right\rVert_2 \\
  &= \left\lVert -Q\left( T -zI\right)^{-1} Q^H \right\rVert_2 \\
  &= \left\lVert \left( T-zI \right)^{-1}\right\rVert_2 \\
  &= \sigma_\text{max}\left( \left( T-zI \right)^{-1} \right).
\end{align*}
Letting \(\lambda_\text{max}(\cdot)\) denote the largest eigenvalue of
a Hermitian matrix,
\begin{align}
  \left\lVert \left( zI-A \right)^{-1} \right\rVert_2
  &= \sqrt{\lambda_\text{max}\left( \left( T-zI \right)^{-H} \left( T-zI \right)^{-1} \right)}.
\end{align}
A Krylov eigensolver like the Lanczos algorithm can estimate the
largest eigenvalue of \(\left( T-zI \right)^{-H} \left( T-zI
\right)^{-1}\) by repeatedly applying it to a vector. Each matrix
product can be computed with two triangular solves, for a total cost
of \(O(n^2)\) flops. The algorithm is outlined below:
\begin{algorithm}[H]
  \caption{Van Loan/Lui algorithm}
  \label{algorithm:van loan/lui}
  \begin{algorithmic}
    \Function{VanLoanLui}{$A, z$}

@@

\appendix
\section{Generalized Multi-Shift Triangular Solve}
The multi-shift triangular solve problem can be generalized and
applied to compute generalized eigenvectors. Given \(m\times m\) upper
triangular matrices \(U\) and \(V\), right-hand side vectors
\(b_1,\cdots,b_n\), and shifts \(\lambda_1,\cdots,\lambda_n\), we seek
solution vectors \(x_1,\cdots,x_n\) such that
\begin{align}
  \left( U - \lambda_j V \right) x_j &= b_j .
\end{align}
This problem can be solved with small modifications to Algorithm
\ref{algorithm:multi-shift trsm}:
\begin{algorithm}[H]
  \caption{Generalized multi-shift triangular solve with blocked back substitution}
  \label{algorithm:generalized multi-shift trsm}
  \begin{algorithmic}
    \Procedure{GeneralizedMultiShiftTrsm}{$U, V, \lambda, B $}
    \Comment \parbox[t]{0.22\linewidth}{\(B\) is overwritten with \(X\)}

    \State \Call{Trsv}{$U', B(\mathcal{I}_1,j)$} 
    \Comment \texttt{xTRSV}

    \State \( B(\mathcal{I}_0,:) := B(\mathcal{I}_0,:) - U(\mathcal{I}_0,\mathcal{I}_1) * B(\mathcal{I}_1,:) \)
    \Comment \texttt{xGEMM}

    \EndProcedure
  \end{algorithmic}
\end{algorithm}
\noindent
Making similar changes to Algorithm \ref{algorithm:safe multi-shift
  trsm} yields a robust algorithm:
\begin{algorithm}[H]
  \caption{Safe, generalized multi-shift triangular solve with
    safeguarded, blocked back substitution }
  \label{algorithm:safe generalized multi-shift trsm}
  \begin{algorithmic}
    \Procedure{SafeGeneralizedMultiShiftTrsm}{$U, V, \lambda, B, s $}
    \Comment \parbox[t]{0.13\linewidth}{\(B\) is overwritten with \(X\)}

    \State \(U' := U(\mathcal{I}_1, \mathcal{I}_1) - \lambda(j) * V(\mathcal{I}_1, \mathcal{I}_1)\) \Comment\texttt{xAXPY}

    \State \Call{SafeTrsv}{$U', B(\mathcal{I}_1,j), t$} 
    \Comment Diagonal block step

    \State \(B(\mathcal{I}_0,j) := t*B(\mathcal{I}_0,j)\) \Comment \texttt{xSCAL}

    \State \(B(\mathcal{I}_2,j) := t*B(\mathcal{I}_2,j)\) \Comment \texttt{xSCAL}

    \State \(C(:,j) := \lambda(j) * B(\mathcal{I}_1,j)\) \Comment \texttt{xSCAL}

    \State \(G(j) := G(j) + \sum_{k\in \mathcal{I}_1} \left( \left\lVert U(\mathcal{I}_0,k) \right\rVert_{\infty} + \left| \lambda(j)\right|  {\left\lVert V(\mathcal{I}_0,k)\right\rVert}_\infty \right) \left\lVert B(\mathcal{I}_1,j) \right\rVert_\infty \)

    \State \(B(:,j) := t * B(:,j)\) \Comment \texttt{xSCAL}

    \State \( B(\mathcal{I}_0,:) := B(\mathcal{I}_0,:) - U(\mathcal{I}_0,\mathcal{I}_1) * B(\mathcal{I}_1,:) \)
    \Comment \texttt{xGEMM}

    \State \( B(\mathcal{I}_0,:) := B(\mathcal{I}_0,:) + V(\mathcal{I}_0,\mathcal{I}_1) * C \)
    \Comment \texttt{xGEMM}

    \EndProcedure
  \end{algorithmic}
\end{algorithm}
\noindent
This routine can be used to solve the generalized eigenvalue problem:
\begin{algorithm}[H]
  \caption{Generalized, triangular and generalized, general
    eigensolvers}
  \label{algorithm:eig}
  \begin{algorithmic}

    \Function{GeneralizedTriangEig}{$T,S$}

    \State \(\lambda := \text{diag}(T) ./ \text{diag}(S)\)

    \State \(Z := -T + S * \text{diag}(\lambda)\)

    \State \Call{SafeGeneralizedMultiShiftTrsm}{$T,S, \lambda, Z, s$}

    \Function{GeneralizedEig}{$A$}

    \State Compute generalized Schur decomposition \(A=QTP^H, B=QSP^H\)

    \State \(\lambda, Z := \) \Call{GeneralizedTriangEig}{$T,S$}

    \State \(X := P Z\) \Comment \texttt{xGEMM}

  \end{algorithmic}
\end{algorithm}
